High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales

### 2021-03-04 inst/manuscript/manuscript.Rmd
Edits to Discussion nearly complete

- name: Su Ye
  affilnum: '1'

@@

- name: Justus Muhando
  affilnum: '2'

- name: Ryan Avery
  affilnum: '3'

- name: Kelly Caylor
  affilnum: '3-5'

@@

- affilnum: 1
  affil: Graduate School of Geography, Clark University, 950 Main Street, Worcester, MA 01610 USA

- affilnum: 2
  affil: SpatialCollective

@@

##### Mapping assignments
Labellers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Figure \ref{fig:labeller}), where they are presented with a white target box representing the randomly selected grid cell, a set of digitizing tools, and different image backdrops, including true and false color renderings of the the growing season and off-growing season PlanetScope composites, and several virtual globe basemaps. Following a set of pre-defined digitizing rules (see SI), the labeller uses the polygon drawing tool to digitize the boundaries of all crop fields intersecting the target grid cell that are visible within the PlanetScope overlays. To aid with interpretation, the labeller can toggle between the PlanetScope renderings and the basemaps to help form a judgement about what constitutes a field. The labeller assigns each polygon a field class category (e.g. annual cropland, tree crop, fallow), and upon completing all fields submits the assignment to the database. In cases where the target grid cell does not contain any fields, the labeller simply submits the assignment to mark it complete. The labeller is then directed to the next available assignment from a different labelling task.

@@

#### Classification
Once the features from the training sites are extracted into `RasterFrames`, these are combined with their corresponding labels and passed to the machine learning classifier, a `SparkMLlib` implementation of Random Forests [@BreimanRandomForests2001]. For this study, the model was trained with a balanced sample and a tree depth of 15 and total tree number of 80, which initial testing showed to provide a reasonable balance between computational time and performance.

@@

## Applying the system to map Ghana
To create a map of Ghana, we divided the country into 16 distinct AOIs, which were developed by grouping together each one degree cell fully contained within Ghana with the tiles belonging to any adjacent degree cell that overlapped neighboring countries (Figure \ref{fig:aois}A). The exception was AOI 16, which consisted of the four degree cells along Ghana's southern coast. The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$), A separate active learning and segmentation process was run for each of these AOIs.

To collect the initial randomized samples for model training, we grouped the AOIs into three clusters: a northern cluster comprising the 6 northernmost AOIs (Cluster 1), a central to southeastern cluster (Cluster 2) consisting of the 3 middle (AOIs 7-9) and 2 southeastern AOIs (12 and 15), and a southwestern cluster (Cluster 3) made up of the forest zone AOIs (10, 11, 13, 14, 16). Within each cluster, we randomly selected and labelled 500 grid cells, which provided relatively large initial training samples for these agro-ecologically similar regions, while helping to minimize the overall amount of labelling effort. In addition to these samples, we randomly selected and labelled 100 grid cells within each AOI to provide a validation sample.

@@

To evaluate the performance of the system, we undertook several different analyses.

@@

After training of models with the initial randomly selected label sets, the active learning process was run for 3 iterations for 12 of the 16 AOIs, resulting in `r typicaln` labels per AOI (see Figure S3A for sample distributions). AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (>83%) and showed little subsequent improvement. The models for these two AOIs were thus trained with `r aoi10n` -  `r aoi14n` samples. AOI 15 was run for 4 iterations (`r aoi15n` samples), while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI). In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S3A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`. The total number of unique training and validation sites across the country were `r ttrain` and `r tval`, respectively.

@@

These results demonstrate a capability for mapping the characteristics of smallholder-dominated cropping systems at high spatial resolution, annual time steps, and national scales. The resulting maps provide an updated and more granular view of the distribution and extent of arable croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017]. This prior work found that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana in 2015, whereas our 2018 maps have cropland cover of 16.1-23.2% (Table \ref{tab:mapaccuracy}), and our map reference sample-based estimates finds `r segcipct[1]`-`r rfcipct[1]`% cover. Our results thus suggest that Ghana's arable cropland is less than previously estimated, but the difference may be attributable to the earlier maps using a more expansive definition of cropland than we did (active to recently active fields).

In addition to the more detailed update of cropland extent, our maps also provide new information on the size and number of fields in Ghana (Figures \ref{fig:mainmap}, S12). Prior efforts to map such agricultural characteristics have often focused on estimating farm rather than field size using census datasets [@jayneAfricaChangingFarm2016; @LowderNumberSizeDistribution2016; @SambergSubnationaldistributionaverage2016; @VonBraunSmallscalefarmersliberalized2004]. Efforts to map fields in small-scale agricultural systems have either been field-based [@CarlettoFactartifactimpact2013; @CarlettoGuesstimatesGPStimatesLand2015] or remote sensing studies over relatively small areas [e.g. @forkuorIntegrationOpticalSynthetic2014] (**include Persello**). The most notable efforts to estimate field sizes were two global initiatives that used crowdsourced volunteers to categorize five broad field size classes based on their interpretations of imagery sampled from virtual globes [@lesivEstimatingGlobalDistribution2019; @FritzMappingglobalcropland2015]. The most recent of these [@lesivEstimatingGlobalDistribution2019] provided country-specific results for Ghana (n = 263) that return an average field size of `r round(lesiv_est, 2)` ha (calculated as the weighted mean of the class counts and the middle values of the four smallest size classes and the lower bound of the largest size class). This estimate exceeds our Ghana-wide average segment size (3.92 ha; Table \ref{tab:sizentab}), but is closer to the mean segment size (`r aoi_nofor` ha) across AOIs 1-9, 12, and 15, which is where most of their sample appears to have been collected. However, our bias-corrected estimates of 1.73 (Ghana-wide) and `r aoi_nofor_adj` (AOIs 1-9, 12, and 15) ha are much smaller.

## Sources of error
Although the maps generated by our system provide valuable new information, they nevertheless contain substantial errors. The overall map accuracies are good (86.7-88%, Table \ref{tab:mapaccuracy}), which is at the boundary of what could be considered achievable map accuracy [@ElmesAccountingtrainingdata2020], given the level of confidence in the map reference sample (only 85% of reference labels were assigned with confidence). However, accuracies for the cropland class were between 62 (Producer's) and 67 (User's) percent in the per-pixel map across the country, indicating that both commission and omission errors were above 30% for this class. The field boundary has lower omission error (Producer's accuracy = 79%), but that comes at the expense of increased false positives (User's accuracy = 58.2%). These accuracies are near the middle to upper ranges of those reported for the cropland class in other large-area mapping studies [@hackmanNewLandcoverMaps2017; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].

The patterns of cropland-class accuracies varied by zone, which largely align with agroecozones and thus somewhat reflective of the inherent challenge each poses for mapping. Producer's accuracies for both maps were highest in the two northern zones (1 and 2), which are primarily savanna agroecoregions (Figure S9), and lowest in zones 3 and 4, which are either forest or coastal savannas. User's accuracies followed a similar pattern, with the exception of Zone 3, which had the highest User's accuracy, albeit from a very small sample. Confining the samples to specific agroecozones provides further insight into error patterns (Table S4). The coastal savannas in the southeast had the highest Producer's accuracy but lowest User's accuracy for the per-pixel map, presumably due to the numerous areas of high density cropland and low woody cover in surrounding uncultivated areas, which promotes commission error. The northern savannas had the best balance between omission and commission error, and had the highest overall User's accuracy. The transition zone had a very low Producer's accuracy (21%), which likely reflects the fact that this region, which also has high cropland density, was divided between several AOIs for mapping.

The relatively low accuracies for the cropland class can be attributed to several important factors. The first of these was the overall mapping extent combined with the resolution of the analysis. Given the goal of country-scale mapping at high resolution, we found that the computational cost required us to use a relatively small and simple set of features and less than the recommended tree number and depth [@MaxwellImplementationmachinelearningclassification2018] in our Random Forest implementation. Previous work shows that RandomForest achieves much better performance on small-scale croplands when trained on a large number of randomly generated features [@Debatsgeneralizedcomputervision2016], which is a commonality with deep learning. However, applying such a large feature set within the extent of our AOIs was intractable, as the computing time would have been several-fold larger than the typical 10-14 hours required from training through prediction on 800 CPUs, for a single iteration. This left the model less able to discern cropland from surrounding savanna vegetation with bare patches or sparse grass cover, which was common in most AOIs. A larger feature set with more variable textures filters could have helped distinguish such areas from crop fields, helping to particularly reduce commission error.

Another major limiting factor was the inherent difficulty of the labelling task. Our system was designed to minimize the error inherent in labelling, but determining croplands from non-croplands over these system is difficult and requires the evaluation of multiple image sources and considerable judgement by each labeller. Interpretation is particularly hard over areas where the background savanna vegetation and croplands have similar reflectance during the dry season, which is a particular problem in AOIs 2 and 3. Such changes in scene circumstances require labellers to adopt different heuristics, for example by delineating the field boundary on the growing season rather than the dry season image. Smaller field sizes also complicated the labelling task, becoming increasingly indistinct in the ~4 m PlanetScope composites. The difficulty of labelling is reflected in the magnitude of the Bayesian Risk metrics (Figure S5), and by labeller accuracies that averaged 71% (Figure S4B). The design of our system allowed us to measure how much label quality impacts model performance. Although prior work [@mellorExploringIssuesTraining2015; @rodriguez-galianoAssessmentEffectivenessRandom2012] has reported that RandomForest is particularly robust to training data error, we found its impact to be substantial (Figure \ref{fig:trainingimpact}), suggesting that the greatest single investment towards improved accuracy lies in improving training data quality.

Image quality was another issue, although primarily in the forest zone (Zone 4) AOIs, where frequent cloud cover reduced the number of available images in all seasons, resulting in composites with more brightness artifacts and blur (Figure \ref{fig:imqual}). This impacted labellers' abilities to discern arable fields, and doubtless affected model predictions. There is little to be done to mitigate these errors, short of confining imagery to the less cloudy dry season, which may further undermine model performance--prior work shows that imagery from multiple seasons is important for improving classifier performance [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019]. Composite quality could be improved by integrating imagery from the same seasons over multiple years, but this would undermine the goal of developing annual maps, while the dynamism of the croplands would confound the ability to delineate field boundaries.

The final major source of error arose from the segmentation process. The vectorized maps had high commission errors that are attributable to the RandomForest predictions. Uncertainty in that model resulted in substantial numbers of pixels just above the 0.5 classification threshold, meaning a number of segments over highly uncertain areas were retained as fields. A more accurate classifier would reduce such errors. Another possible solution could be to apply a locally varying classification threshold to distinguish croplands from non-cropland [e.g. @waldnerDeepLearningEdge2020]. Over-merging was another source of error in the segmentation algorithm. This led in some areas to over-estimation of field sizes and often irregular shapes, particularly in high density croplands (e.g. in AOIs 2 and 8; Figure \ref{fig:mainmap}) where adjacent fields have boundaries that are indistinct in the PlanetScope imagery. Minimizing or preventing merging would help in such cases, although it would result in  over-segmentation and thus would likely underestimate field sizes. However, over-segmentation can help to identify important within-field distinctions (e.g. different soil types), which can be useful for other applications, such as filtering pixels for crop type mapping or yield estimation [e.g. @EstesComparingmechanisticempirical2013; @EstesProjectedclimateimpacts2013]

## Key system features
Despite the sources of errors mentioned above, the design of the system proved effective in mitigating error, resulting in a higher overall accuracy than would have otherwise been possible. The label accuracy assessment and consensus labelling methods appeared to be the most effective tools for error mitigation. Label accuracy assessment allowed us to quantify the impact of trainnig data error on model performance, in keeping with recommended best practices [@ElmesAccountingtrainingdata2020]. Label quality proved to have a large impact on model performance (Figure \ref{fig:trainingimpact}), and consensus labeling was effective in reducing individual digitization errors and between-labeller differences in interpretation, improving label, and thus map, quality. Worker-specific label accuracy measures also helped to improve the overall quality of the consensus labels, by enabling higher weight to be placed on labels that were more likely to be accurate during the merging process, rather than giving equal weight to potentially less accurate labels. The ability to select the most accurate individual labels for a site also allowed us to develop independent estimates of field size to which measures of confidence can be attached (Figure S4B), which we could in turn use to bias-correct map-based size class estimates (Table \ref{tab:sizentab}). The ability to select the most accurate label will also help when different training strategies are needed, such as when training a model to perform instance segmentation [@carvalhoInstanceSegmentationLarge2021], in which case the field identities have to be preserved.

```{r, message = FALSE, warning = FALSE, echo=FALSE}
prdelta <- iteration_metrics %>% 
  filter(Metric %in% c("FPR", "TPR")) %>% 
  group_by(Iteration, Metric) %>% 
  summarize(Score = mean(Score)) %>% 
  filter(Iteration %in% c(0, 3)) %>% 
  tidyr::pivot_wider(names_from = Iteration, values_from = Score) %>% 
  mutate(rate = round((`0` - `3`) / `0` * 100, 1))
```
The active learning approach helped to improve overall model performance relative to a randomized approach to training site selection. Although we only compared active to random site selection within three AOIs, the average gains across all three AOIs and three performance metrics demonstrated the effectiveness of the active approach, most notably by the gains in the F1 score (Figure S7), a combined metric of commission and omission error. The gains in accuracy and AUC were smaller. For accuracy, this was due to the larger drop in the active learning metric after iteration 1 in AOI 15, which proved one of the hardest AOIs to map. The plateau in AUC gains at 0.5% relative to random training reflects the fact that each iteration tended to reduce both the false and true positive rates simultaneously. Although the false positive rate overall declined (an average of `r prdelta$rate[1]`% between Iterations 0 and 3), the reductions in true positive rate (averaging `r prdelta$rate[2]`%) offset some of the improved predictive power.

The gain due to active learning were substantially less than those reported by Debats et al (2017), who found that active site selection improved classifier performance by `r round((0.45-0.35) / 0.35 * 100)` over randomized site selection after one iteration, and after 45 iterations by `r round((0.69-0.64) / 0.64 * 100)`%. These larger gains likely reflect the fact that this evaluation began from the smallest possible training sample--one image chip--and tested the effect of adding a single additional chip each iteration. In contrast, we started with a large initial random sample (500) and added 100 samples per iteration, thus our models received significant training at the outset, making the marginal gain due to each new sample smaller. This suggests that we could have benefitted by reducing the size of the initial random training sample while increasing the size of actively selected samples, not only in terms of larger accuracy gains, but also in terms of reducing--by as much as 62%, per Debats et al (2017)--the overall size of the training sample needed to achieve a given level of performance.

The detail, large extent, and temporal precision of the maps we created by virtue of our system's intentional design around PlanetScope imagery, which currently provides the only wall-to-wall, sub-5 meter imaging capability [@McCabefutureEarthobservation2017] with daily frequency. Daily frequency is important for creating multiple seasonal composites during a single agricultural year in cloudy regions. Although unquantified, our compositing technique may be more effective at providing complete image coverage during a season than the "best-on-top" approach\footnote{https://assets.planet.com/products/basemap/planet-basemaps-product-specifications.pdf} used by Planet to develop their analytic basemap, which anecdotally appears to contains substantial residual cloud cover during the growing season in southern Ghana. Although free and providing higher radiometric quality and depth, and of sufficiently high resolution (10 m) to accurately classify small-scale agricultural systems [e.g. @Defournyrealtimeagriculturemonitoring2019; @kernerRapidResponseCrop2020], Sentinel-2's 5-day revist may be too infrequent to provide enough clear pixels to create a cloud-free growing season composite in the southern half of Ghana. Sentinel-1 is not affected by the same problem, but labelling fields in a more coarsely resolved radar images is difficult.

## Lingering questions
Several potential issues not answered by our assessment merit further exploration. These include the correspondence between image- and ground-collected labels. We believe that such comparisons will in most circumstances reveal unresolvable discrepancies. The highly dynamic nature of the systems in question means that even narrow temporal discrepancies between ground and image-collected labels can produce substantial differences. These discrepancies could be further exacerbated by differences in an understanding of what constitutes a field, which might vary on the ground depending on the farmer or the data collecter's collectors judgement. On top of this, field-collected boundaries are typically highly geographically constrained and limited in size, meaning that any comparison might suggest that the relationship is more noisy than it actually is. These factors suggest that ground versus image label differences would not necessarily indicate how far image-based labellers were from the "truth". Nevertheless, a comparison against ground data can help to assess how accurately image-collected labels capture the typical size of fields, and thus merits further investigation for this reason.

@@

## Next steps
The maps presented here represent a version 1 product that is freely available to use **(links)**, along with its underlying code. These data were developed according to the recommended best practices for training and assessing error in machine learning models [@ElmesAccountingtrainingdata2020]. In their current form, the maps may be useful for a variety of research applications. For example, analyzing the  distributions of values in the probability maps may provide additional insight into the relative extents of active versus fallow croplands [@Tongforgottenlanduse2020]. However, use of these data, particularly for decision-making processes (e.g. cropped area estimates), should be careful to account for the reported errors [@OlofssonGoodpracticesestimating2014; @StehmanKeyissuesrigorous2019].

>>>>

- name: Su Ye
  affilnum: '1,2'

@@

- name: Justus Muhando
  affilnum: '3'

- name: Ryan Avery
  affilnum: '4'

- name: Kelly Caylor
  affilnum: '4-6'

@@

- affilnum: 1
  affil: Graduate School of Geography, Clark University, 950 Main Street, Worcester, MA, USA

- affilnum: 2
  affil: Department of Natural Resources and the Environment, University of Connecticut, Storrs, CT, USA

- affilnum: 2
  affil: SpatialCollective, Nairobi, Kenya

@@

##### Mapping assignments
Labellers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Figure \ref{fig:labeller}), where they are presented with a white target box representing the randomly selected grid cell, a set of digitizing tools, and different image backdrops, including true and false color renderings of the the growing season and off-growing season PlanetScope composites, and several virtual globe basemaps. Following a set of pre-defined digitizing rules (see SI), the labeller uses the polygon drawing tool to digitize the boundaries of all crop fields intersecting the target grid cell that are visible within the PlanetScope overlays. To aid with interpretation, the labeller can toggle between the PlanetScope renderings and the basemaps to help form a judgement about what constitutes a field. The labeller assigns each polygon a class category (e.g. annual cropland), and upon completing all fields submits the assignment to the database. In cases where the target grid cell does not contain any fields, the labeller simply submits the assignment to mark it complete. The labeller is then directed to the next available assignment from a different labelling task.

@@

#### Classification

Once the features from the training sites are extracted into `RasterFrames`, these are combined with their corresponding labels and passed to the machine learning classifier, a `SparkMLlib` implementation of Random Forests [@BreimanRandomForests2001]. For this study, the model was trained with a balanced sample and a tree depth of 15 and total tree number of 80, which initial testing showed to provide a reasonable balance between computational time and performance.

@@

## Applying the system to map Ghana
To create a map of Ghana, we divided the country into 16 distinct AOIs, which were developed by grouping together each one degree cell fully contained within Ghana with the tiles belonging to any adjacent degree cell that overlapped neighboring countries (Figure \ref{fig:aois}A). The exception was AOI 16, which consisted of the four degree cells along Ghana's southern coast. The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$), A separate active learning and segmentation process was run for each of these AOIs. 
To collect the initial randomized samples for model training, we grouped the AOIs into three clusters: a northern cluster comprising the 6 northernmost AOIs (Cluster 1), a central to southeastern cluster (Cluster 2) consisting of the 3 middle (AOIs 7-9) and 2 southeastern AOIs (12 and 15), and a southwestern cluster (Cluster 3) made up of the forest zone AOIs (10, 11, 13, 14, 16). Within each cluster, we randomly selected and labelled 500 grid cells, which provided relatively large initial training samples for these agro-ecologically similar regions, while helping to minimize the overall amount of labelling effort. In addition to these samples, we randomly selected and labelled 100 grid cells within each AOI to provide a validation sample.

@@

During the collection of training and validation samples, labellers were tasked to only digitize active or recently active crop fields, avoiding tree crops, and fallow or potentially abandoned fields (see SI for the digitizing rules).

To evaluate the performance of the system, we performed several analyses described in sections 2.3.1-4.

@@

After training of models with the initial randomly selected label sets, the active learning process was run for 3 iterations for 12 of the 16 AOIs, resulting in `r typicaln` labels per AOI. AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (>83%) and showed little subsequent improvement. The models for these two AOIs were thus trained with `r aoi10n` -  `r aoi14n` samples. AOI 15 was run for 4 iterations (`r aoi15n` samples), while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI). In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S3A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`. Labels collected during the active learning iterations showed distinct patterns in several AOIs, which often falling along ecotones, such as the boundaries between agroecozones (see Figure S3A). The total number of unique training and validation sites across the country were `r ttrain` and `r tval`, respectively.

@@

These results demonstrate a capability for mapping the characteristics of smallholder-dominated cropping systems at high spatial resolution, annual time steps, and national scales. The resulting maps provide an updated and more granular view of the distribution and extent of arable croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017]. This prior work found that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana in 2015, whereas our 2018 maps have cropland cover of 16.1-23.2% (Table \ref{tab:mapaccuracy}), and our map reference sample-based estimates finds `r segcipct[1]`-`r rfcipct[1]`% cover. Our results thus suggest that Ghana's arable cropland is less than previously estimated, but the difference is perhaps attributable to our use of a cropland definition that excluded longer fallows and abandoned fields, which in some regions can account for over half of the total area that could be counted as cropland [@Tongforgottenlanduse2020].

In addition to the more detailed update of cropland extent, our maps also provide new information on the size and number of fields in Ghana (Figures \ref{fig:mainmap}, S11-12). Previous work to estimate such agricultural characteristics have often focused on farm, rather than field, size using census data [@jayneAfricaChangingFarm2016; @LowderNumberSizeDistribution2016; @SambergSubnationaldistributionaverage2016; @VonBraunSmallscalefarmersliberalized2004]. Efforts to map field boundaries in smallholder-dominated agricultural systems have either used *in situ* data collection [@CarlettoFactartifactimpact2013; @CarlettoGuesstimatesGPStimatesLand2015] or remote sensing studies over relatively small [e.g. @forkuorIntegrationOpticalSynthetic2014; @perselloDelineationAgriculturalFields2019] or discontiguous [@Estesplatformcrowdsourcingcreation2016] areas. The most extensive studies to date used crowdsourced volunteers to classify fields into broad size classes, based on their interpretations of imagery sampled from high resolution virtual globes [@lesivEstimatingGlobalDistribution2019; @FritzMappingglobalcropland2015]. Those efforts included country-specific results for Ghana (n = 263), which can be converted into an average field size estimate of `r round(lesiv_est, 2)` ha$\footnote{Obtained by calculating the weighted mean from the count of the five size classes and the mean of the hectare range provided for the four smallest size classes, and the lower bound of the size range provided the largest size class. Data sourced from Table S3 in Lesiv et al. 2019.}$. This estimate exceeds our Ghana-wide average segment size (3.92 ha; Table \ref{tab:sizentab}), but is closer to the mean (`r aoi_nofor` ha) within AOIs 1-9, 12, and 15, which is where most of the crowdsourced sample appears to have been collected. However, our bias-corrected estimates of 1.73 (Ghana-wide) and `r aoi_nofor_adj` (AOIs 1-9, 12, and 15) ha were much smaller.

## Map accuracy and key sources of error
Although the maps generated by our system provide valuable new information, they nevertheless contain substantial errors. The overall map accuracies are decent (86.7-88%, Table \ref{tab:mapaccuracy}), and are near the boundary of what might be considered *achievable* map accuracy [@ElmesAccountingtrainingdata2020], given the inherent uncertainty in the map reference sample, our best estimate of the "truth", in which we have roughly 85% confidence. However, accuracies for the cropland class were much lower, falling between 62 (Producer's) to 67 (User's) percent country-wide for the per-pixel map (Table \ref{tab:mapaccuracy}), meaning the model produced substantial commission and omission errors for this class. The segmented boundary maps had fewer omission errors (Producer's accuracy = 79%), but higher false positive errors (User's accuracy = 58.2%). These accuracies are near the middle to upper ranges of those reported for the cropland class in other large-area mapping studies [@hackmanNewLandcoverMaps2017; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].

The patterns of cropland-class accuracies varied by zone. These zones largely align, albeit with some discrepancies, with the country's agroecoregions, thus the accuracy patterns may be partially attributed to some regions being harder to map than others. Producer's accuracies for both maps were highest in the two northern zones (1 and 2), which are primarily savannas (Figure S9), and lowest in zones 3 and 4, which are comprised of forest or coastal savannas. User's accuracies followed a similar pattern, with the exception of Zone 3, which had the highest User's accuracy, albeit from a very small sample. Aligning the reference samples more precisely with agroecozone boundaries (Figure S9B) provides further insight into error patterns (Table S4). Coastal savannas in the southeast had the highest Producer's accuracy but lowest User's accuracy for the per-pixel map, presumably because this region's numerous areas of high density cropland, combined with low woody cover in surrounding uncultivated areas, helped to promote commission error. Maps in the two northern savanna agroecoregions had the best balance between omission and commission error, and had the highest overall User's accuracy. The transitional (between forest and savanna) agroecozone had a very low Producer's accuracy (21%), which likely reflects the fact that it was divided between several AOIs for mapping (Figure S3), within which it typically covered a smaller share of area relative to the other agroecoregions. This likely caused insufficient representation of this AEZ in training samples, particularly in AOIs 10 and 11 (Figure S3B).

Beyond the errors linked to regional differences, several other important factors contributed to reducing the accuracies in the cropland class. The first of these stems from the overall mapping extent and the high resolution of the data. Given the goal of developing a country-scale map at high resolution, the attendant data volume required us to use a relatively small set of image features and less than the recommended tree number and depth [@MaxwellImplementationmachinelearningclassification2018] in our Random Forests implementation, in order to limit computational costs. Previous work found that Random Forests achieves much better performance on small-scale croplands when trained on a much larger number of features [@Debatsgeneralizedcomputervision2016]. However, applying such a large feature set within the extent of our AOIs was intractable, as the computing time would have been several-fold larger than the ~4-8 hours of runtime on 800 CPUs required for a single active learning iteration, followed by ~10-14 hours for prediction. This reduced the skill of the model, particularly when it came to differentiating cropland from adjacent bare patches or natural vegetation with sparse herbaceous cover, which were common in many AOIs.

The inherent difficulty of the labelling task was another major limiting factor. Our system was designed to minimize the error inherent in labelling, but determining croplands from non-croplands in these agricultural systems can be a difficult task. Labellers have to evaluate multiple image sources and to rely heavily on judgement, which inevitably leads to errors. Interpretation is particularly hard where the background savanna vegetation and croplands have similar reflectance during the dry season, which is a particular problem in AOIs 2 and 3. Smaller field sizes also complicate labelling, as these become increasingly indistinct in the ~4 m PlanetScope composites. The difficulty of labelling is reflected in the magnitude of the Bayesian Risk metrics (Figure S5), and by the average score achieved by each labeller against our training reference dataset  (71%; Figure S4B). Although prior work [@mellorExploringIssuesTraining2015; @rodriguez-galianoAssessmentEffectivenessRandom2012] found that Random Forests are robust to label errors, we found that they have substantial impact (Figure \ref{fig:trainingimpact}), which suggest that simply improving label quality may be one of the single most important investments towards improved model accuracy.

Image quality was another issue, although primarily in the forested AOIs, where frequent cloud cover reduced the number of available images in all seasons, resulting in composites with more brightness artifacts and blur (Figure \ref{fig:imqual}). This impacted labellers' abilities to discern arable fields, and doubtless affected model predictions. There is little to be done to mitigate these errors, short of confining imagery to the less cloudy dry season, which may further undermine model performance, given the importance of multi-temporal imagery for cropland classification [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019]. Composite quality could be improved by using imagery from the same seasons over multiple years, but this would undermine the goal of developing annual maps, while the dynamism of the croplands would blur field boundaries within the imagery.

The final major source of error arose from the segmentation process. The vectorized maps had high commission errors caused by uncertainties in the Random Forests predictions. Model uncertainty led to many pixels over non-crop areas with probabilities straddling the 0.5 classification threshold. Segments that intersected such areas were retained as fields when the average probability of intersecting pixels exceeded 0.5. A more accurate classifier would reduce such errors, or the application of a locally varying classification threshold [e.g. @waldnerDeepLearningEdge2020]. Over-merging was another source of error in the segmentation algorithm, which in some areas led to overestimated field sizes and irregular shapes, particularly in high density croplands (e.g. in AOIs 2 and 8; Figure \ref{fig:mainmap}) where the boundaries between adjacent fields are often indistinct in the PlanetScope imagery. Minimizing or preventing merging would help in such cases, although could would result in the opposite problem, over-segmentation, and thereby underestimate field size.

## Error mitigation features
Despite the error sources mentioned above, several features of the system proved effective in mitigating error, leading to a higher overall accuracy than would have otherwise been possible. Label accuracy assessment and consensus labelling appeared to be the most effective error mitigation tools. Label accuracy measures allowed us to quantify the substantial impact of label error on model performance (Figure \ref{fig:trainingimpact}), while consensus labels substantially reduced individual labelling errors, resulting in maps that were more accurate than they would have been had we used individually generated labels. Labeller-specific accuracy measures also helped to improve the quality of the consensus labels, by placing higher weight on labels more likely to be accurate during the merging process, rather than giving equal weight to potentially less accurate labels. The ability to select the most accurate individual labels for a site also allowed us to develop independent estimates of field size to which measures of confidence can be attached (Figure S4B), which we were in turn able to correct estimates of field sizes and numbers (Table \ref{tab:sizentab}).

```{r, message = FALSE, warning = FALSE, echo=FALSE}
prdelta <- iteration_metrics %>% 
  filter(Metric %in% c("FPR", "TPR")) %>% 
  group_by(Iteration, Metric) %>% 
  summarize(Score = mean(Score)) %>% 
  filter(Iteration %in% c(0, 3)) %>% 
  tidyr::pivot_wider(names_from = Iteration, values_from = Score) %>% 
  mutate(rate = round((`0` - `3`) / `0` * 100, 1))
```
The active learning approach helped to improve overall model performance relative to randomized training site selection, confirming previous findings reported by @DebatsIntegratingactivelearning2017. Although the performance gains relative to randomized model training that we observed were substantially less than in the previous study, which reported `r round((0.45-0.35) / 0.35 * 100)`% higher model performance after one iteration, and `r round((0.69-0.64) / 0.64 * 100)`% higher after 45 iterations [@DebatsIntegratingactivelearning2017], those comparisons were made from a low initial base, beginning with a single training site with one more site added each iteration. In contrast, our large initial random sample (500) meant that our models were substantially trained before they were exposed to actively selected labels, thereby diluting their impact on performance. Nevertheless, the higher average performance of the active approach across three performance metrics demonstrated its effectiveness. Most notable were the larger gains in the F1 score (Figure S7), a balanced performance metric. The gains in accuracy and AUC were smaller. For accuracy, the reduced advantage was primarily due to active learning being outperformed by randomized training after the first iteration in AOI 15, which proved one of the hardest AOIs to both map and label. Active learning likely resulted in the selection of sites that were harder to label than randomly selected ones, leading to more label error, and thus initially lower model accuracy. However, this deficit was overcome by the 5th iteration. The plateau in AUC gains at 0.5% better than randomized training reflects the findings that active learning reduced both the false and true positive rates, the two inputs to AUC. Although the false positive rate declined by `r prdelta$rate[1]`% between Iterations 0 and 3, the `r prdelta$rate[2]`% reduction in the true positive rate offset some of these gains, and the effect of this tradeoff may have been more pronounced in the threshold independent AUC.

The detail, temporal precision, and large extent of these maps was enabled by the system's ability to process PlanetScope data, which is currently the only source of sub-5 meter imagery with daily coverage [@McCabefutureEarthobservation2017]. Daily revisits are necessary for creating seasonal composites within a single agricultural year over cloudy areas. The compositing technique we developed to work with PlanetScope was able to provide complete image coverage over the entire country for both seasons. Although Sentinel-2 is free, has better radiometric quality, and has sufficiently high resolution (10 m) to accurately classify small-scale agricultural systems [e.g. @Defournyrealtimeagriculturemonitoring2019; @kernerRapidResponseCrop2020], its 5-day interval is too infrequent to generate sufficiently cloud-free composites during the growing season over southern Ghana. Sentinel-1 is not affected by the same problem, but labelling fields in more coarsely resolved radar images is difficult.

## Lingering questions
Several potential issues not answered by our assessment merit further exploration. One of these is the degree of correspondence between image- and ground-collected labels. However, such comparisons will likely reveal unresolvable discrepancies in most circumstances. The highly dynamic nature of the systems in question means that even narrow temporal discrepancies between ground and image-collected labels may show substantial differences. These discrepancies could be further exacerbated by differences in an understanding of what constitutes a field, which might vary on the ground depending on the farmer or the data collecter's collectors judgement. On top of this, field-collected boundaries are typically highly geographically constrained and limited in size, meaning that any comparison might suggest that the relationship is more noisy than it actually is. These factors suggest that ground versus image label differences would not necessarily indicate how far image-based labellers were from the "truth". Nevertheless, a comparison against ground data can help to assess how accurately image-collected labels capture the typical size of fields, and thus merits further investigation for this reason.

@@

## Next steps
The maps presented here represent a version 1 product that is freely available to use, along with its underlying code. These data were developed according to the recommended best practices for training and assessing error in machine learning models [@ElmesAccountingtrainingdata2020]. In their current form, the maps may be useful for a variety of research applications. For example, analyzing the distributions of values in the probability maps may provide additional insight into the relative extents of active versus fallow croplands [@Tongforgottenlanduse2020]. However, use of these data, particularly for decision-making processes (e.g. cropped area estimates), should be careful to account for the reported errors [@OlofssonGoodpracticesestimating2014; @StehmanKeyissuesrigorous2019].
